\documentclass[11pt,doublespace]{unhthesis}


% \includeonly{
% sections/Introduction,
% sections/SatelliteAttitudeModeling,
% sections/TableSat1A,
% sections/TableSatAttitudeDynamicsSoftware,
% sections/StateError,
% sections/Estimators,
% sections/Controllers,
% sections/ObserverBasedControls,
% sections/SoftwareProgression,
% sections/TSatPy,
% sections/Conclusions,
% sections/FutureWork,
% sections/EquationSummary,
% }

\usepackage{custom}
\renewcommand\listingscaption{Snippet}

\hyphenation{gno-mon-ly}


\begin{document}

\title{Development of a Modular Application for Analyzing Observer-Based Control Systems for NASA's Spin-Stabilized MMS Mission Spacecraft}
\author{Daniel R. Couture}
\prevdegrees{B.S. Mechanical Engineering, University of New Hampshire, 2001}
\major{Mechanical Engineering}
\degree{Master of Science}
\degreemonth{May}
\degreeyear{2014}
\thesisdate{March 15, 2014}
\DOCUMENTtype{THESIS}
\Documenttype{Thesis}
\documenttype{thesis}
\maketitle

\abstractcontents{This thesis utilizes an experimental table top satellite prototype (NASA MMS TableSat IA) to span two main efforts.  The first is to create a physical model of a satellite for NASA's Magnetospheric MultiScale (MMS) Mission in order to validate and compare varied gyroless attitude determination and control systems (ADCS) techniques.  The ADCS maintains the TableSat constant 3 rpm rotation, prevents boom oscillations, and detects and corrects for unwanted nutation.  The second effort is a software system that can be used to implement theoretical and experimental models of rotational motion.  The system provides near ``real-time'' feedback of the system's states, and allows for ``on-the-fly'' modification to control parameters.  The system is specifically designed such that any control systems engineer is able to customize and extend its functionality without substantial programming expertise.}

% \makeannouncement

\copyrightyear{2014}
\makecopyright

\supervisor{Dr. May-Win L. Thein}{Associate Professor of Mechanical Engineering}
\committee{Dr. Barry Fussell}{Professor of Mechanical Engineering}
\committee{Dr. Philip J. Hatcher}{Professor of Computer Science}
\makeapproval

\begin{dedication}
\begin{flushleft}
To Dominic and Ian.  May you always make time in your life for a little curiosity.

To Heather.  This has been an incredible experience, and I can't imagine doing it without you.
\end{flushleft}
\end{dedication}

\begin{acknowledgments}
Since starting my Master's degree almost 13 years ago, there have been a large and wide group of people that have guided and supported my journey:

Professor Thein most of all who has stuck with me far beyond what sanity should have allowed.

My fellow cast and crew of the TableSat IA team for building the experimental platform on which this research was performed.

Professors Fussell and Hatcher for volunteering precious time to serve on my thesis committee.

The MMS Attitude Control System Group of the Goddard Space Flight Center, specifically: Josephine San, Dean Chai, and David Mangus.

Tracey Harvey for the omnipresent cheer and smoothing out all the scheduling and paperwork issues complicated by being a part time student.

The NH Space Grant Consortium and NASA MMS Grants for the research support.

My esteemed colleagues at the Newmarket School District that not only are the supportive second family, but established the foundation of my academic career.

My parents for stressing the importance of learning.
\end{acknowledgments}

\begin{singlespace}
  \tableofcontents
  \addcontentsline{toc}{section}{LIST OF TABLES}
  \listoftables
  \addcontentsline{toc}{section}{LIST OF FIGURES}
  \listoffigures
\end{singlespace}

\begin{abstractpage}
\end{abstractpage}


\include{sections/Introduction}
\include{sections/SatelliteAttitudeModeling}
\include{sections/TableSat1A}
\include{sections/TableSatAttitudeDynamicsSoftware}
\include{sections/StateError}
\include{sections/Estimators}
\include{sections/Controllers}
\include{sections/ObserverBasedControls}
\include{sections/SoftwareProgression}
\include{sections/TSatPy}
\include{sections/Conclusions}
\include{sections/FutureWork}

\include{sections/Biblio}

\newpage
\vspace*{100mm}
\begin{center}
  {\huge \bf APPENDICES}
\end{center}

\appendix
\pagebreak
\begin{singlespace}
  \include{sections/EquationSummary}
  \include{sections/TSatPySamples}
  \include{sections/TSatPySource}
  \include{sections/NSSSource}
\end{singlespace}




% \munfig{Credit: http://mms.gsfc.nasa.gov}{2d_field_lines}{pretty/2d_earth_mag_field_lines.png}{0.4}


% \section{Mission}

% Science questions to answer \cite{mms_website}
%     What determines when reconnection starts and how fast it proceeds?
%     What is the structure of the diffusion region?
%     How do the plasmas and magnetic fields disconnect and reconnect in the diffusion regions?
%     What role do the electrons play in facilitating reconnection?
%     What is the role of turbulence in the reconnection process?
%     How does reconnection lead to the acceleration of particles to high energies?




% % \munfig{Credit: Southwest Research Institute}{mms_spacecraft_formation}{pretty/mms_spacecraft_formation.jpg}{0.4}

% % \munfig{Credit: http://earthobservatory.nasa.gov}{northern_lights_iss}{pretty/Northern-lights-ISS.jpg}{0.4}

% % \munfig{Credit: http://mms.gsfc.nasa.gov}{tetrahedron}{pretty/mms_dayside_tetrahedral_thumb.png}{0.4}

% % \munfig{Credit: http://mms.gsfc.nasa.gov}{2d_field_lines}{pretty/2d_earth_mag_field_lines.png}{0.4}
% % \munfig{Credit: http://mms.gsfc.nasa.gov}{instrument_deck_webview}{pretty/instrument_deck_webview.jpg}{0.4}
% % \munfig{Credit: http://mms.gsfc.nasa.gov}{observatory_dimensions1_webview}{pretty/observatory_dimensions1_webview.jpg}{0.4}
% % \munfig{Credit: http://mms.gsfc.nasa.gov}{observatory_interfaces_webview}{pretty/observatory_interfaces_webview.jpg}{0.4}
% % \munfig{Credit: http://mms.gsfc.nasa.gov}{observatory_webview}{pretty/observatory_webview.jpg}{0.4}

% % \munfig{Credits: ASRC Research and Technology/Barbara Lambert}{assembly}{pretty/713494main2_MMS-Observatory2-assembled-670.jpg}{0.4}

\end{document}