

\chapter{System Iterations}
\label{chap:SystemIterations}

There was a bit of trial and error and multiple full system rewrites of the TSat ADCS.

\TODO{Elaborate on the self imposed criteria for the code. Emphasize the stress on use for outreach programs and reuse in subsequent TSat style projects.
}

\section{Control Loop in Matlab Simulink}
\label{sec:ControlLoopinSimulink}

Slight variation of the ``open loop'' control system used in part of Melissa Vess' TableSat research.  Huge problem with fault tolerance.  The format chosen for relaying messages resulted in an unreliable exchange of data between the base station (laptop) and the TableSat.  These error were compounded with Matlab Simulink's inability to handle the unexpected lack of message data.

\TODO{More detail here on the Simulink side.  Specify pros of this design.  Include screen shot of simulink diagram for a P controller.}

\section{TSat Message Center - Matlab Graphical User Interface (GUI)}
\label{sec:TSatMessageCenter}

This was the first full rework of the control system.  The focus of this design was to eliminate the use of Matlab Simulink diagrams, and directly control the messages transmitted and received.

\TODO{Elaborate on the failures of the design.  Inability to handle two tasks at one.  State persistence.  No visualization.  }

\section{Runtime Control and Analysis - Advanced Matlab GUI}
\label{sec:RuntimeControlandAnalysis}

\section{TSatPy - Python Package}
\label{sec:TSatPyPythonPackage}

Chapter \ref{chap:TSatPy} will cover the final version of the UNH TableSat 1A's ADCS design in more detail.

